Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-07-25 surfaces-write-up/surface-write-up.tex
surfaces: minor intro rewording, implement restate

\usepackage{amsmath}
\usepackage{amsxtra}
\usepackage{amscd}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{eucal}
\usepackage[all]{xy}
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{mathrsfs}
\usepackage{subfiles}
%\usepackage{mathpazo} not a huge fan
\usepackage{euler}
\usepackage{hyperref}
\usepackage{color}
\usepackage{longtable}
\usepackage{float}
\usepackage{caption}

@@

In part, this work is motivated by a recent paper by
O'Dorney \cite{dorney:canonical} who gives similar descriptions
of canonical rings for general $\bq$-divisors on $\bp^1$. This
study also has connections to the papers by Voight--Zureick-Brown 
\cite{vzb:stacky} and Landesman--Ruhm--Zhang \cite{lrz:spin-cring} on log 
spin canonical rings of particular $\bq$ divisors on stacky curves which 
provide bounds on rings of modular forms of Fuchsian groups.

@@

\todo{Aaron: This pararaph was already stated above. Can we remove it?}
This work extends the study of canonical rings of
$\bq$-divisors on curves, as done by O'Dorney \cite{dorney:canonical},
Voight--Zureick-Brown \cite{vzb:stacky}, and Landesman--Ruhm--Zhang
\cite{lrz:spin-cring}, to general projective space and Hirzebruch
surfaces.

\todo{more! Aaron: I think this is plenty of introduction}

\ssec{Main Results}

\todo{put main results! use restate, maybe?}
\todo{Aaron: Some of these theorem statements have been copied from older versions of the statements
and not updated to reflect the difference. I agree we should use Theorem-restate, but we should make
sure to copy the ones stated below, and not copy these ones}
\begin{thm}
\label{thm:proj-effective-intro}
Let $X = \bp^m$ with $\{x_0, \ldots, x_m\}$ a coordinate system for
$\bp^m$. Let $D = \sum_{i = 1}^{n} \alpha_i C_i$, with $\alpha_i =
\frac{p_i}{q_i} \in \bq$ and each $C_i$ a hypersurface of
$\bp^m$, be an effective divisor on $\bp^m$.

@@

\begin{thm}
\label{thm:proj-generators-relations-intro}
Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\mathbb{P}^m$.  Let
$f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $C_i = v(f_i)$
for all $i$; suppose $\deg(f_i) = a_i$. 
Then $R_D$ is generated in degree at most 
\[
	\sum_{i=1}^n \ell_i a_i
\]
with relations generated in degree at most
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{thm}

@@

\begin{thm}
\label{thm:hirz-generators-relations}
Let $D = \sum_{i=1}^n \alpha_i C_i$ for curves $C_i$ on $\hirz_k$.  
Let $C_i = V(f_i)$ where $f_i \in \sco(a_i, b_i)$; further 
suppose $f_1$ and $f_2$ are independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$.
Then $R_D$ is generated in degree less than 
\[
	\sigma = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i +
	\sum_{\substack{i	\in \Tp(D) \\	j \in \Tm(D)}} (\bida_i
	\bidb_j - \bida_j \bidb_i)\ell_{i,j}
\]

\noindent
with relations generated in degree less than 
\[
	\max(2 \sigma, \sigma + \tau)
\]

\noindent
where
\[
	\tau = \sigma
	+ \max \left( \max_{i\in \Te}(\ell_i \gcd(a_i, b_i)), \max_{
	\substack{i \in	\Tp \\ j\in \Tm}}(\ell_{i,j} (\bida_i \bidb_j
	- \bida_j \bidb_i))	\right).
\]
\end{thm}

\todo{Include summary of what each of the sections do.}
\todo{Incorperate the following section into the pm section}
\section{Lemmas for Generation and Relations}
\label{sec:lemmas}

@@

Combining the above results, we obtain the following bound on the
degrees of generators and relations of $R_D$.

\begin{thm}
\label{thm:proj-generators-relations}
Let $D = \sum_{i=1}^n \alpha_i D_i \in \bq \otimes \di \bp^m$. Let
$f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that
$D_i = V(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$. 
Then $R_D$ is generated in degree at most 
\[
	\sum_{i=1}^n \ell_i a_i
\]

\noindent
with relations generated in degree at most
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}
	(	\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{thm}

@@

@@

Combining these results, we get the following theorem:

\begin{thm}
\label{thm:hirz-generators-relations}
Let $D = \sum_{i=1}^n \alpha_i D_i \in \bq \otimes \di \hirz_k.$ Let $D_i =
V(f_i)$ where $f_i \in \sco(a_i, b_i)$; further suppose $f_1$ and $f_2$ are
independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$.
Then $R_D$ is generated in degrees at most
\[
	\rho = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i +
	\sum_{\substack{i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j
	- \bida_j \bidb_i) \ell_{i, j}
\]

\noindent
where
\[
	\tau = \sigma
	+ \max \left( \max_{i\in \Te} \bigl(\ell_i \gcd(a_i, b_i) \bigr),
	\max_{\substack{i \in \Tp \\ j \in \Tm}} \bigl(\ell_{i, j}
	(\bida_i \bidb_j - \bida_j \bidb_i) \bigr) \right).
\]
\end{thm}
\begin{proof}
The generation degree bound is as stated in Lemma \ref{lem:hirz-generators}.  By Proposition \ref{lem:composite-map}, the degree of generation of $\ker(\phi)$ is at most the maximum of the generation degrees of $\ker(\chi)$ and $\ker(\psi)$, giving us the desired relations bound.
\end{proof}

\todo{Aaron: Shouldn't $\tau$ not include $\sigma$?  Peter: Fixed}

\todo{Aaron: I know exactly what is going on, but please spell out the proof in 2 lines for completeness.\\
Peter: Better?}

>>>>

\usepackage{amsmath}
\usepackage{amsxtra}
\usepackage{amscd}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{eucal}
\usepackage[all]{xy}
\usepackage{graphicx}
\usepackage{tikz-cd}
\usepackage{mathrsfs}
\usepackage{subfiles}
%\usepackage{mathpazo} not a huge fan
\usepackage{euler}
\usepackage{hyperref}
\usepackage{color}
\usepackage{longtable}
\usepackage{float}
\usepackage{caption}
\usepackage{thmtools}
\usepackage{thm-restate}

@@

In part, this work is motivated by a recent paper by
O'Dorney \cite{dorney:canonical}, which gives similar descriptions
of canonical rings for general $\bq$-divisors on $\bp^1$. This
study also has connections to the papers by Voight--Zureick-Brown 
\cite{vzb:stacky} and Landesman--Ruhm--Zhang \cite{lrz:spin-cring}
on log spin canonical rings on stacky curves which provide bounds
on rings of modular forms of Fuchsian groups through the study
of particular $\bq$-divisors on curves.

@@

\ssec{Main Results}
\todo{Aaron: Some of these theorem statements have been copied from older versions of the statements
and not updated to reflect the difference. I agree we should use Theorem-restate, but we should make
sure to copy the ones stated below, and not copy these ones. Robin: thm-restate only allows you to restate earlier theorems and not later ones, so the main thms must appear numbered as 1.-}
\begin{thm}
\label{thm:proj-effective-intro}
Let $X = \bp^m$ with $\{x_0, \ldots, x_m\}$ a coordinate system for
$\bp^m$. Let $D = \sum_{i = 1}^{n} \alpha_i C_i$, with $\alpha_i =
\frac{p_i}{q_i} \in \bq$ and each $C_i$ a hypersurface of
$\bp^m$, be an effective divisor on $\bp^m$.

@@

\begin{restatable}{thm}{restateproj}
\label{thm:proj-generators-relations}
Let $D = \sum_{i=1}^n \alpha_i D_i \in \bq \otimes \di \bp^m$. Let
$f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that
$D_i = V(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$. 
Then $R_D$ is generated in degree at most 
\[
	\sum_{i=1}^n \ell_i a_i
\]

\noindent
with relations generated in degree at most
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}
	(	\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{restatable}

@@

\begin{restatable}{thm}{hirzrestate}
\label{thm:hirz-generators-relations}
Let $D = \sum_{i=1}^n \alpha_i D_i \in \bq \otimes \di \hirz_k.$ Let $D_i =
V(f_i)$ where $f_i \in \sco(a_i, b_i)$; further suppose $f_1$ and $f_2$ are
independent linear
polynomials in $x_0$ and $x_1$ and $f_3$ and $f_4$ are independent
linear polynomials in $y_0$ and $y_1$.
Then $R_D$ is generated in degrees at most
\[
	\rho = \sum_{i\in \Te(D)} \gcd(\bida_i, \bidb_i)\ell_i +
	\sum_{\substack{i \in \Tp(D) \\	j \in \Tm(D)}} (\bida_i \bidb_j
	- \bida_j \bidb_i) \ell_{i, j}
\]

\noindent
where
\[
	\tau = \sigma
	+ \max \left( \max_{i\in \Te} \bigl(\ell_i \gcd(a_i, b_i) \bigr),
	\max_{\substack{i \in \Tp \\ j \in \Tm}} \bigl(\ell_{i, j}
	(\bida_i \bidb_j - \bida_j \bidb_i) \bigr) \right).
\]
\end{restatable}

\todo{Include summary of what each of the sections do.}
\todo{Incorporate the following section into the $\bp^{m}$ section}

\section{Lemmas for Generation and Relations}
\label{sec:lemmas}

@@

By combining these results, we get our main theorem bounding
the generator and relation degrees of the canonical ring of
$\bq$-divisors on projective space.

\restateproj*

@@

\todo{say something about ghost points?}

@@

By combining these results, we get our main theorem bounding
the generator and relation degrees of the canonical ring of
$\bq$-divisors on Hirzebruch surfaces.

\hirzrestate*

\begin{proof}
The generation degree bound is as stated in Lemma \ref{lem:hirz-generators}.  By Proposition \ref{lem:composite-map}, the degree of generation of $\ker(\phi)$ is at most the maximum of the generation degrees of $\ker(\chi)$ and $\ker(\psi)$, giving us the desired relations bound.
\end{proof}

\todo{add something about ghost points?}
\todo{Aaron: I know exactly what is going on, but please spell out the proof in 2 lines for completeness.\\
Peter: Better?}
